---
title: "Homework 3"
author: "Kai Suzuki"
date: "2025-06-11"
format: 
  html:
    toc: true
    toc-depth: 12
---

[Link to Repository](https://github.com/kai-s-suzuki/ENVS-193DS_spring-2025_final)

```{r packages, message=FALSE}
library(tidyverse)
library(here)
library(janitor)
```

# Problem 1. Research writing

## a. 
In part 1, my co-worker used a linear model, likely a test finding the Pearson correlation coefficient based on the continuous values. In part 2, a one-way ANOVA test was likely used, as they are looking for one continuous variable being different between multiple sources. 

## b. 
In part 2, they could have shown us the variances of each source and ran a Tukey's honest significance test. Showing the variances between the different sources can explain whether the test assumed equal variances or not; if the variances are similar--within four times of each other as a gut test, it would change the perception of the information. A Tukey's HSD test run after would then tell us which sources were different from each other. 

## c. 

# Problem 2. Data visualization

## a. 
```{r}
sst <- read_csv("../data/SST_update2023.csv")
```

```{r}
# data wrangling
sst_clean <- sst |> # starting with the sst data frame
  filter(year(date) >=2018) |> # filtering to only include dates with year at or after 2018
  mutate(year = year(date), # adding a year column
        month = month(date, label = TRUE), # adding a month column, with letters (Jan) instead of numbers (1)
        month = as_factor(month)) |> # making months a factor with 12 variables
    group_by(year, month) |> # grouping by month and year to allow data to be summarized
  summarize(mean_monthly_sst = round(mean(temp), 1)) # finding the mean sea surface temperatures of every measurement taken in each month in the data frame, rounded to one decimal placestr

sst_clean |>
  ungroup() |>
slice_sample(n = 5)

str(sst_clean)
```

