---
title: "Homework 3"
author: "Kai Suzuki"
date: "2025-06-11"
format: 
  html:
    toc: true
    toc-depth: 12
---

[Link to Repository](https://github.com/kai-s-suzuki/ENVS-193DS_spring-2025_final)

```{r packages, message=FALSE}
library(tidyverse)
library(here)
library(janitor)
library(DHARMa)
library(MuMIn)
```

# Problem 1. Research writing

## a. 
In part 1, my co-worker used a linear model, likely a test finding the Pearson correlation coefficient based on the continuous values. In part 2, a one-way ANOVA test was likely used, as they are looking for one continuous variable being different between multiple sources. 

## b. 
In part 2, they could have shown us the variances of each source and ran a Tukey's honest significance test. Showing the variances between the different sources can explain whether the test assumed equal variances or not; if the variances are similar--within four times of each other as a gut test, it would change the perception of the information. A Tukey's HSD test run after would then tell us which sources were different from each other. 

## c. 

# Problem 2. Data visualization

## a. 
```{r}
sst <- read_csv("../data/SST_update2023.csv")
```

```{r message = FALSE}
# data wrangling
sst_clean <- sst |> # starting with the sst data frame
  filter(year(date) >=2018) |> # filtering to only include dates with year at or after 2018
  mutate(year = year(date), # adding a year column
         year = as_factor(year),
        month = month(date, label = TRUE), # adding a month column, with letters (Jan) instead of numbers (1)
        month = as_factor(month)) |> # making months a factor with 12 variables
    group_by(year, month) |> # grouping by month and year to allow data to be summarized
  summarize(mean_monthly_sst = round(mean(temp), 1)) # finding the mean sea surface temperatures of every measurement taken in each month in the data frame, rounded to one decimal place

sst_clean |> # starting with the sst_clean data frame
  ungroup() |> # ungrouping so that sliced samples are not taken for each year
slice_sample(n = 5) # slicing 5 samples from any five months

str(sst_clean) # looking at underlying data and factors
```

## b.
```{r warning = FALSE}
ggplot(data = sst_clean, # starting with the sst_clean dataframe
    aes(x = month, # setting x axis as month
         y = mean_monthly_sst, # setting y axis as mean sea surface temperature
        color = year, # coloring by year
        group = year)) + # grouping by year (connects points for each year)
  geom_point() + # adding each data point
  geom_line() + # connecting points
  scale_color_manual(values = c("2018" = "#86dfa2", # coloring 2018 as the lightest green
                                "2019" = "#71d290", # coloring 2019 as a slightly darker green than 2018
                                "2020" = "#55bb76", # coloring 2020 as a slightly darker green than 2019
                                "2021" = "#42ab64", # coloring 2021 as a slightly darker green than 2020
                                "2022" = "#197236", # coloring 2022 as a slightly darker green than 2021
                                "2023" = "#105827")) + # coloring 2023 as the darkest green
  labs(x = "Month", # labeling the x axis as "Month"
       y = "Mean monthly sea surface temperature (°C)", # labeling the y axis
       color = "Year") + # relabeling the legend to "Year"
  theme_bw() + # using black and white ggplot theme
  theme(legend.position = c(0.1, 0.7), # moving legend position to inside graph
        panel.grid.major = element_blank(), # removing major lines, which are vertical lines and some of the horizontal
        panel.grid.minor = element_blank() # removing minor lines, which are horizontal
  ) + 
  scale_y_continuous( # changing the scale of the graph
    breaks = seq(14, 20, by = 2), # matching given visualization, with ticks at 14 through 20
    limits = c(12.4,19.65)) # manually setting the min and max of the y-axis, adjusting so that '20' is aligned with the top of the graph and the lowest point is appropriately displayed and settled at the bottom
```

# Problem 3. Data analysis

```{r}
nest_boxes <- read_csv("../data/occdist.csv") # reading in data from article
```


## a. 

Biologically, a '0' in either the 'sp', 'cs', or 'tm' columns, means that the box was either unoccupied, or a different bird species occupied the box; a '0' in the 'e' column means that the box was occupied by any bird. 
A '1' in either the 'sp', 'cs', or 'tm' columns means that the box was occupied by one or more of that bird species; a '1' in the 'e' column means that the box was unoccupied by any bird. 

## b. 

In the context of this study, Swift Parrots are the target species for occupants of these bird boxes, as they are picky in terms of nesting locations compared to the Common Starling and Tree Martins. Neither species is particularly dominant over the other, but the Swift Parrot has suffered the most due to its selective behavior and requires the most attention in terms of conservation. 

## c. 

The 2016 'season' represents the first year of the study that coincided with a tree blooming event, where the boxes were deployed for the first time at the location. The 2019 'season' represents another tree blooming event, with the boxes left with uncertainty of whether the parrots would return to the same location for nesting. 

## d. 

| Model number | Season | Distance to Forest Edge |        Model Description     |  
|:------------:|:------:|:-----------------------:|:----------------------------:|
| 0            |        |                         |  no predictors (null model)  |
| 1            |  X     |            X            |  all predictors (sat model)  | 
| 2            |  X     |                         |      Season only             |   
| 3            |        |            X            | Distance to Forest Edge only |    

## e. 

```{r}
nest_boxes_clean <- nest_boxes |> # starting with the nest_boxes dataframe
  clean_names() |> # cleaning names
  select(box_occupant, edge_distance, season, sp) # selecting specific rows
```

```{r}
model0 <- glm(sp ~ 1, # linear model without predictors
              data = nest_boxes_clean, # using the clean dataframe
              family = binomial)
model1 <- glm(sp ~ season + edge_distance, # linear model of parrot occupancy predicted by season and distance
             data = nest_boxes_clean, # using the clean dataframe
             family = binomial)
model2 <- glm(sp ~ season, # linear model of parrot occupancy predicted by season
             data = nest_boxes_clean, # using the clean dataframe
             family = binomial)
model3 <- glm(sp ~ edge_distance, # linear model of parrot occupancy predicted by distance
             data = nest_boxes_clean, # using the clean dataframe
             family = binomial)
```

```{r}
plot(simulateResiduals(model0))
plot(simulateResiduals(model1))
plot(simulateResiduals(model2))
plot(simulateResiduals(model3))
```

```{r}
summary(model0)
summary(model1)
summary(model2)
summary(model3)
```

